Notes and code examples to accompany "If the Null Fits, You Must Omit

### 2025-12-09 GroupPolarizationStatmod/code_examples.qmd
finally documenting well to avoid backtracking through codebase every time

### Purpose: demonstrate gp-stats codebase

@@

Change directories to go to the `paper` directory where this document lives:

```         
cd gp-statmod/GroupPolarizationStatmod/paper
```

@@

![**Mean group opinions pre- and post-discussion from Moscovici & Zavalloni
(1969).**](images/clipboard-1443885641.png){width="575"}

@@

@@

print(head(fwer_tbl))
```

@@

This is reported for a variety of effect-detection thresholds, $\delta^*$, which the function name implies are "significance" values. When I first wrote this function
which are really effect-detection thresholds—so the code slightly misuses the term "significance", which typically refers only to a p-value based effect-detection threshold in the *null-hypothesis significance testing* approach. I updated the prose in the article to reflect this distinction, but left the code as-is. I again call the function with the defaults for illustration.

@@

fdr_over_thresholds <- calc_fdr_vs_significance(
  probit_fits_dir, 
)
```

@@

>>>>

## Purpose: demonstrate gp-stats codebase

@@

Change directories to go to the `gp-statmod/GroupPolarizationStatmod` directory where this document lives:

```         
cd gp-statmod/GroupPolarizationStatmod
```

@@

![**Mean group opinions pre- and post-discussion from Moscovici & Zavalloni
(1969).**](code_examples_images/clipboard-1443885641.png){width="575"}

@@

The first step in our _generative_ model of false detections is to _generate_ simulated 
experimental data from a mere agreement null-polarization model. Then we can calculate how frequently an effect size was inferred that exceeded the effect-detection threshold.  simulated null-model data our simulated null-models.

## Simulate Experimental Data

@@

print(head(fwer_tbl, n = 10))
```

@@

The `calc_fdr_vs_significance` function calculates the false-positive rate ($\alpha$) and false detection rate ($\mathrm{FDR}$) reported for a variety of effect-detection thresholds, $d^*$, as we call them in the paper. When I first wrote the code, though, I was thinking about detection thresholds as synonymous with "significance values". So that's why the function name, `calc_fdr_vs_significance`, includes the word `significance`, which is a stand-in for the maximally descriptive "effect-detection thresholds" terminology. This seemed fine to keep things simple and not change the name everywhere in the code. But it was right to change it for the paper, since *significance* is only understood in a p-value based  *null-hypothesis significance testing* approach, where a hypothesis is confirmed if the p-value is less than some particular threshold. I updated the prose in the article to reflect this distinction, but left the code as-is.

I again call the function with the defaults specified, for illustration. Remember the `significance_vals` map to _effect-detection thresholds_ ($d^*$) in the paper.

@@

fdr_over_thresholds <- calc_fdr_vs_significance(
  probit_fits_dir = probit_fits_dir,  # defined in above code block
  base_rate = 0.1, power = 0.8, significance_vals = c(0.2, 0.5, 0.8)
)

print(head(fdr_over_thresholds, n=10))
```
To see `SigVal = 0.8` (ie $d^* = 0.8$), check the `tail` of the table:
```{r}
print(tail(fdr_over_thresholds, n = 10))
```

@@

# Analyses

## _Identify_ Invalid Replications: All But 5% Were Invalidated

## _Evaluate_ Experimental Designs: ~70% Chance a Polarization Detection was Erroneous

## _Specify_ Effect-Detection Thresholds: Spec Experiments—Don't Just Design Them—and Avoid Informed Guesswork

# References
